---
title: "3CT"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: '2022-12-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}
setwd("~/Documents/Systemic ESYT1 project locomotion, behavior, survival, anatomical data/Behavioral paradigms data processing/Behavior analysis for paper/Analysis R scripts/3CT/")

library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans, DescTools)

df <- read_excel("3CT results.xlsx")

```

Assign genotypes
```{r}
df$Genotype <- 0

df$Genotype <- ifelse(df$ID==702, 'SOD1', df$Genotype)
df$Genotype <- ifelse(df$ID==730, 'SOD1', df$Genotype)
df$Genotype <- ifelse(df$ID==734, 'SOD1', df$Genotype)
df$Genotype <- ifelse(df$ID==738, 'SOD1', df$Genotype)
df$Genotype <- ifelse(df$ID==625, 'SOD1', df$Genotype) 
df$Genotype <- ifelse(df$ID==627, 'SOD1', df$Genotype) 
df$Genotype <- ifelse(df$ID==615, 'SOD1', df$Genotype) 
df$Genotype <- ifelse(df$ID==788, 'SOD1', df$Genotype) 
df$Genotype <- ifelse(df$ID==700, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==692, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==706, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==708, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==553, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==555, 'WT', df$Genotype)
df$Genotype <- ifelse(df$ID==694, 'En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==698, 'En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==696, 'En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==561, 'En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==736, 'SOD1En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==589, 'SOD1En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==603, 'SOD1En1cre', df$Genotype)
df$Genotype <- ifelse(df$ID==607, 'SOD1En1cre', df$Genotype)

```

Removal of mouse 625 and 627 from dataset as they were wrongly injected.
```{r}
df<-df[!(df$ID==625 | df$ID==627),] 

```

Divide dataset up into habituation, social, and novelty phases

```{r}
habituation_cond <-  df %>% filter(Condition == "Hab")
social_cond<-  df %>% filter(Condition == "Social")
novelty_cond<-  df %>% filter(Condition == "Novel")
```

Calculating preferences for zones for the different conditions - as described in Baronio 2015 
#https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0116363&type=printable 

```{r}
#This can be used for looking at preference for either side during habituation if needed
habituation_cond <- habituation_cond %>% mutate(social_preference_relative=((`In zone SocialZone cumulative s`-`In zone NoveltyZone cumulative s`)/(`In zone NoveltyZone cumulative s`+`In zone SocialZone cumulative s`)))

#calculating preference for social side during social condition if needed
social_cond <- social_cond %>% mutate(social_preference_relative=((`In zone SocialZone cumulative s`-`In zone NoveltyZone cumulative s`)/(`In zone NoveltyZone cumulative s`+`In zone SocialZone cumulative s`)))

#calculating preference for novelty side during novelty condition
novelty_cond <- novelty_cond %>% mutate(novelty_preference_relative=((`In zone NoveltyZone cumulative s`-`In zone SocialZone cumulative s`)/(`In zone SocialZone cumulative s`+`In zone NoveltyZone cumulative s`)))

```


Test genotype differences - novelty preference

```{r}
#refactor genotype so SOD1 becomes control for Dunnet's post-hoc
novelty_cond$Genotype <- factor(novelty_cond$Genotype, levels = c("SOD1","SOD1En1cre", "WT", "En1cre"))

###Relative preference
anova_novel_preference_relative <- aov(novelty_preference_relative ~ Genotype, data = novelty_cond)
summary(anova_novel_preference_relative)

#Dunnet's post hoc
DunnettTest(x=novelty_cond$novelty_preference_relative, g=novelty_cond$Genotype)

```

##Plots

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice

```{r}
color_palette = c("#5B1A18","#D67236","#FD6467", "#F1BB7B")
```

Preparing the ggsignif function to use Dunnet's p-values
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}

```

```{r}

Dunnet<- glht(anova_novel_preference_relative, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

#Re-ordering so the experimental groups appear in order of control level
novelty_cond$Genotype <- factor(novelty_cond$Genotype, levels = c("WT", "En1cre","SOD1", "SOD1En1cre"))

ggplot(data= novelty_cond, aes(x=Genotype, y=novelty_preference_relative)) +
  ylab("Social Novelty Preference Index")+
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.4, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 
  theme(text = element_text(family = "Arial"))+
  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  geom_jitter(position = position_jitter(width = .2),alpha=0.6) +
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(family = "Arial",
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(0.65,0.55,0.45)
  )
```

