---
title: "Lateral analysis"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading data

```{r}
library(pacman)
p_load(tidyverse, ggsignif, multcomp, broom, emmeans, DescTools)
setwd("~/Documents/Systemic ESYT1 project locomotion, behavior, survival, anatomical data/Locomotion analysis cleaned/") 
df<- read.table("statistics.csv", sep=",", header=T)


```
Assign time point

```{r}
df2 <- df
df2$time_point <- 0
df2$time_point <- ifelse(grepl('p49_', df2$name), 49, df2$time_point)
df2$time_point <- ifelse(grepl('p56_', df2$name), 56, df2$time_point)
df2$time_point <- ifelse(grepl('p63_', df2$name), 63, df2$time_point)
df2$time_point <- ifelse(grepl('p70_', df2$name), 70, df2$time_point)
df2$time_point <- ifelse(grepl('p77_', df2$name), 77, df2$time_point)
df2$time_point <- ifelse(grepl('p84_', df2$name), 84, df2$time_point)
df2$time_point <- ifelse(grepl('p91_', df2$name), 91, df2$time_point)
df2$time_point <- ifelse(grepl('p98_', df2$name), 98, df2$time_point)
df2$time_point <- ifelse(grepl('p105_', df2$name), 105, df2$time_point)
df2$time_point <- ifelse(grepl('p112_', df2$name), 112, df2$time_point)

```



Give each mouse own number
```{r}

df2$mouse_number <- 0

#SOD1En1cre
df2$mouse_number <- ifelse(grepl('736_', df$name), '736', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('589_', df$name), '589', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('603_', df$name), '603', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('607_', df$name), '607', df2$mouse_number)
#En1re
df2$mouse_number <- ifelse(grepl('561_', df$name), '561', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('694_', df$name), '694', df2$mouse_number) 
df2$mouse_number <- ifelse(grepl('696_', df$name), '696', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('698_', df$name), '698', df2$mouse_number)
#SOD1
df2$mouse_number <- ifelse(grepl('730_', df$name), '730', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('734_', df$name), '734', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('738_', df$name), '738', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('702_', df$name), '702', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('627_', df$name), '627', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('625_', df$name), '625', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('615_', df$name), '615', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('788_', df$name), '788', df2$mouse_number)
#WT
df2$mouse_number <- ifelse(grepl('692_', df$name), '692', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('700_', df$name), '700', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('706_', df$name), '706', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('708_', df$name), '708', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('553_', df$name), '553', df2$mouse_number)
df2$mouse_number <- ifelse(grepl('555_', df$name), '555', df2$mouse_number)
```

Assign genotypes
```{r}
df2$Genotype <- 0


df2$Genotype <- ifelse(df2$mouse_number==702, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==730, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==734, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==738, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==625, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==627, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==615, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==788, 'SOD1', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==700, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==692, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==706, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==708, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==553, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==555, 'WT', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==694, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==698, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==696, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==561, 'En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==736, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==589, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==603, 'SOD1En1cre', df2$Genotype)
df2$Genotype <- ifelse(df2$mouse_number==607, 'SOD1En1cre', df2$Genotype)

```


Removal of mouse 625 and 627 from dataset as they wrongly injected.

```{r}
df2<-df2[!(df2$mouse_number=='625' | df2$mouse_number=='627'),] 
```

Compute mean for each mouse for each timepoint

```{r}
df_means_mouse_by_Genotype <- df2 %>%
	group_by(mouse_number, Genotype, time_point) %>%
  summarise(hip_ang = mean(hip_ang), knee_ang = mean(knee_ang), ankle_ang = mean(ankle_ang), foot_ang = mean(foot_ang)) %>% 
  arrange(Genotype)
```

Create dataset with data from onset of locomotor phenotype

Set color palette for plots

For the plots, colors from the palette 'GrandBudapest1' from the 'wesanderson' package (Karthik Ram and Hadley Wickham (2018). wesanderson: A Wes Anderson Palette Generator. R package
  version 0.3.6. https://CRAN.R-project.org/package=wesanderson) were used, but colors re-ordered so the darkest color was applied to the WT mice
```{r}
#color_palette inspired colorpalette
color_palette = c("#5B1A18","#FD6467",  "#F1BB7B")

```

Preparing the ggsignif function to use Dunnet's p-values for bar plots
```{r}

stars <- function(p) {
  case_when(
    p <= 0.001 ~ "***",
    p <= 0.01  ~ "**",
    p <= 0.05  ~ "*",
    TRUE       ~ "ns"
  )
}


```

Analysis of p112

```{r}
p112 <- subset(df_means_mouse_by_Genotype, time_point==112) #Mean values
p112_all <- subset(df2, time_point==112) #Video triplicates
```

one-way anova at p112
 
```{r}

#Ordering Genotype variables by manipulation so SOD1 becomes 'control' in Dunnet post-hoc tests
p112_all$Genotype <- factor(p112_all$Genotype, levels = c("SOD1","SOD1En1cre", "WT", "En1cre"))

#foot angle
anova_foot_1way <- aov(foot_ang ~ Genotype, data = p112_all)
summary(anova_foot_1way)

DunnettTest(x=p112_all$foot_ang, g=p112_all$Genotype)

#ankle
anova_ankle_1way <- aov(ankle_ang ~ Genotype, data = p112_all)
summary(anova_ankle_1way)

DunnettTest(x=p112_all$ankle_ang, g=p112_all$Genotype)

#knee
anova_knee_1way <- aov(knee_ang ~ Genotype, data = p112_all)
summary(anova_knee_1way)

DunnettTest(x=p112_all$knee_ang, g=p112_all$Genotype)

#hip
anova_hip_1way <- aov(hip_ang ~ Genotype, data = p112_all)
summary(anova_hip_1way)

DunnettTest(x=p112_all$hip_ang, g=p112_all$Genotype)


summary(anova_hip_1way)
```



Bar plots p112_all

```{r}


#Re-ordering so the experimental groups appear in order of control level
p112$Genotype <- factor(p112$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))
p112_all$Genotype <- factor(p112_all$Genotype, levels = c("WT", "En1cre", "SOD1", "SOD1En1cre"))


#foot
Dunnet<- glht(anova_foot_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=foot_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Foot angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(), text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(70,62,66)
  )

#ankle
Dunnet<- glht(anova_ankle_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=ankle_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Ankle angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(47,41,44)
  )

#knee
Dunnet<- glht(anova_knee_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=knee_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+ 

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Knee angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(23,20,21.5)
  )

#hip
Dunnet<- glht(anova_hip_1way, mcp(Genotype = "Dunnet")) %>%tidy() %>% 
  mutate(contrast = str_split(contrast, " - "),
         stars    = stars(adj.p.value))

ggplot(data= p112_all, aes(x=Genotype, y=hip_ang)) +
  geom_bar(stat = "summary",fun=mean, aes(color = Genotype,fill=Genotype), alpha = 0.01, size=1) +
  geom_errorbar(stat="summary", fun.data=mean_se, width = 0.5, position = "dodge")+
  theme(panel.background = element_rect(fill = "white"), axis.line = element_line(color="black"))+

  scale_color_manual(values=color_palette)+
  scale_fill_manual(values=color_palette)+
  ylab("Hip angle")+
    geom_jitter(data=p112, position = position_jitter(width = .2), size=2.2)+
  geom_jitter(data=p112_all,position = position_jitter(width = .2),alpha=0.3)+
    theme(axis.title.x=element_blank(),  text = element_text(size=15),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
      ggsignif::geom_signif(
    comparisons = Dunnet$contrast,
    annotations = Dunnet$stars,
    y_position = c(60,50,55)
  )


```
